---
title: "RMBL_NICHE-QNEST-Data-Clean"
author: "Beth Morrison"
date: "April 9, 2020"
output:
  prettydoc::html_pretty:
    highlight: github
---

# Data

The datasets cleaned in ths code are the QNEST bloom and plant-pollinator interaction data

These data are the "KCAqnest_network_observations_RMBL_2019_RAW.csv" (interaction observations)
"qnest_plant_RMBL_2017_RAW.csv" (qnest bloom data for 2017)
"qnest_plant_RMBL_2016_RAW.csv" (qnest bloom data for 2016)
"qnest_plant_RMBL_2019_RAW.csv" (qnest bloom data for 2018 and 2019)

The files are on drop box and DO NOT HAVE THE `_RAW` SUFFIX -- I added that myself when I downloaded the data to my local as it is my preference to denote raw datasets. But, when you download the data from dropbox you can call them anything you want, just make sure your new name matches the R code in the file name argument of `read.csv`. 

The niche dyanmics analysis that uses the objects created in this code is in the `RMBL_NICHE-Niche-Dynamics-Analysis` file

# Libraries

```{r message = FALSE, warning = FALSE}

library(vegan)
library(bipartite)
library(stringr)
library(vegan)
library(spaa)

```

# Data

Interaction observations

```{r}

raw.ds <- read.csv(file = "KCAqnest_network_observations_RMBL_2019_RAW.csv", head = T, stringsAsFactors = F)

```

Bloom observations

```{r}

t1 <- read.csv(file = "qnest_plant_RMBL_2019_RAW.csv", head = T)
t2 <- read.csv(file = "qnest_plant_RMBL_2017_RAW.csv", head = T)
t3 <- read.csv(file = "qnest_plants_RMBL_2016_RAW.csv", head = T)

t1 <- t1[ ,c(1:6)]
t2 <- t2[ ,c(1:6)]
t3 <- t3[ ,c(1:6)]

colnames(t1) ## date site transect segment plant total_flowers
colnames(t2) ## date site transect segment plant_name total_flowers
colnames(t3) ## date site transect segment plant_name total_flowers

colnames(t1)[5] <- 'plant_name'

t2 <- t2[which(t2$site != ""), ]
length(unique(t2$site))

raw.bloom.ds <- rbind(t3, t2, t1)

raw.bloom.ds$site <- droplevels(raw.bloom.ds$site)

```
## Interaction observations data cleaning

Checking for weird data and typos

```{r}

table(raw.ds$date) ## some are lower than I would think, but is that intentional? hard to say... 
table(raw.ds$site) ## G0thic Town
table(raw.ds$who) ## don't really care I guess
table(raw.ds$transect) ## p4 and p6 have only one, some are blank
table(raw.ds$segment) ## looks good
table(raw.ds$vial_number) ## lots, do I care?
table(raw.ds$gross_ID) ## two blank, thrip? should include? Unknown? should remove, should also group the bombus, no? Given the coarse levels for other groups, bombyliidae and bombilidae, Other Fly and Other fly , Big megachile and Big Megachile
table(raw.ds$plant) ## four blank, Hymenoxys hoopesii, HYmenoxys hoopesii
table(raw.ds$em_morpho) ## not sure if i need this information
table(raw.ds$fine_ID) ## 7778 blank!

```

Updating weird data and typos

```{r}

raw.ds1 <- raw.ds

raw.ds1$site[which(raw.ds1$site == 'GOthic Town')] <- 'Gothic Town'
# raw.ds1$site <- droplevels(raw.ds1$site)

raw.ds1 <- raw.ds1[which(raw.ds1$gross_ID != ""), ]
raw.ds1$gross_ID[which(raw.ds1$gross_ID == "Big Megachile")] <- 'Big megachile'
raw.ds1$gross_ID[which(raw.ds1$gross_ID == levels(raw.ds$gross_ID)[8])] <- 'Bombyliidae'
raw.ds1$gross_ID[which(raw.ds1$gross_ID == "Other Fly")] <- 'Other fly'
raw.ds1$gross_ID[which(raw.ds1$gross_ID == "Other fly ")] <- 'Other fly'
raw.ds1 <- raw.ds1[which(raw.ds1$gross_ID != "Thrip"), ]
raw.ds1 <- raw.ds1[which(raw.ds1$gross_ID != "Unknown"), ]
#raw.ds1$gross_ID <- droplevels(raw.ds1$gross_ID)

raw.ds1 <- raw.ds1[which(raw.ds1$plant != ""), ]
raw.ds1$plant[which(raw.ds1$plant == "HYmenoxys hoopesii")] <- "Hymenoxys hoopesii"
#raw.ds1$plant <- droplevels(raw.ds1$plant)

table(raw.ds1$site) ## looks good
table(raw.ds1$gross_ID) ## looks good
table(raw.ds1$plant) ## looks good too

raw.ds1$betterdate <- as.Date(raw.ds1$date, format = "%m/%d/%y")

nrow(raw.ds1) # 8097 interactions

colnames(raw.ds1)
unique(raw.ds1$betterdate) ## 4 years

```

This is kind of an aside but I pull out all the interactions that just include bombus, incase you ever wanted to look at Bombus exclusively, since currently the bombus data are the most well resolved

```{r}

## still need to group all of the Bombus

raw.ds1$gross_ID <- as.character(raw.ds1$gross_ID)
raw.ds1$gross_ID[which(str_detect(raw.ds1$gross_ID, "Bombus"))] <- "Bombus"
table(raw.ds1$gross_ID) ## okay now all the bombus are one

bombus.ds <- raw.ds[which(str_detect(raw.ds$gross_ID, "Bombus")), ]
#bombus.ds$gross_ID <- droplevels(bombus.ds$gross_ID)
bombus.ds$gross_ID <- as.character(bombus.ds$gross_ID)
bombus.ds <- bombus.ds[which(bombus.ds$plant != ''), ]
#bombus.ds$plant <- droplevels(bombus.ds$plant)
#bombus.ds$site <- droplevels(bombus.ds$site)
bombus.ds$betterdate <- as.Date(bombus.ds$date, format = "%m/%d/%y")

```

## Bloom data cleaning

Looking for weird data and typos

```{r}

table(raw.bloom.ds$date) ## dates look good
table(raw.bloom.ds$site) ## good 
table(raw.bloom.ds$transect) ## some blanks to remove
table(raw.bloom.ds$segment) ## also look good as far as I can tell
sort(names(table(raw.bloom.ds$plant))) ##  
table(raw.bloom.ds$total_flowers) ## nothing immediately shocking but we shall see

```

Fixing weird data and typos

```{r}

raw.bloom.ds$plant <- as.character(raw.bloom.ds$plant)
raw.bloom.ds <- raw.bloom.ds[which(raw.bloom.ds$transect != ""), ]
raw.bloom.ds$plant[which(raw.bloom.ds$plant == "Agoseris  glauca")] <- "Agoseris glauca"
raw.bloom.ds <- raw.bloom.ds[which(raw.bloom.ds$plant != "Antemeria White"), ]
raw.bloom.ds <- raw.bloom.ds[which(raw.bloom.ds$plant != ""), ]
raw.bloom.ds$plant[which(raw.bloom.ds$plant == "Aquilegia coerulea")] <- "Aquilegia caerulea"
raw.bloom.ds <- raw.bloom.ds[which(raw.bloom.ds$plant != "Arnica? "), ]
raw.bloom.ds <- raw.bloom.ds[which(raw.bloom.ds$plant != "Big Yellow Bushy Flowers"), ]
raw.bloom.ds$plant[which(raw.bloom.ds$plant == "Potentilla gracilis")] <- "Potentilla pulcherrima"
raw.bloom.ds$plant[which(raw.bloom.ds$plant == 'Bochera stricta')] <- 'Boechera stricta'
raw.bloom.ds$plant[which(raw.bloom.ds$plant == 'Fringed brassica ')] <- 'Fringed Brassica'
raw.bloom.ds$plant[which(raw.bloom.ds$plant == 'lathyrus leucanthus')] <- 'Lathyrus leucanthus'
raw.bloom.ds$plant[which(raw.bloom.ds$plant == 'Fringed brassica ')] <- 'Fringed Brassica'
raw.bloom.ds$plant[which(raw.bloom.ds$plant == 'Linum Lewisii')] <- 'Linum lewisii'
raw.bloom.ds$plant[which(raw.bloom.ds$plant == 'LInum lewisii')] <- 'Linum lewisii'
raw.bloom.ds$plant[which(raw.bloom.ds$plant == 'Mertensia ciliata ')] <- 'Mertensia ciliata'
raw.bloom.ds$plant[which(raw.bloom.ds$plant == 'penstemon whippleanus')] <- 'Penstemon whippleanus'
raw.bloom.ds$plant[which(raw.bloom.ds$plant == 'Potentilla fruiticosa')] <- 'Potentilla fruticosa'
raw.bloom.ds$plant[which(raw.bloom.ds$plant == 'Erigeron speciosa')] <- 'Erigeron speciosus'
raw.bloom.ds$plant[which(raw.bloom.ds$plant == 'Erigeron speciosus ')] <- 'Erigeron speciosus'
raw.bloom.ds$plant[which(raw.bloom.ds$plant == 'potentilla pulcherrima')] <- 'Potentilla pulcherrima'
raw.bloom.ds <- raw.bloom.ds[which(raw.bloom.ds$plant != 'Snap Dragon Ask Loy!'), ] 
raw.bloom.ds <- raw.bloom.ds[which(raw.bloom.ds$plant != 'Thistle (Need ID)'), ] 
raw.bloom.ds$plant[which(raw.bloom.ds$plant == 'viola praemorsa')] <- 'Viola praemorsa'
raw.bloom.ds$plant[which(raw.bloom.ds$plant == 'Viccia americana')] <- 'Vicia americana'
raw.bloom.ds$plant[which(raw.bloom.ds$plant == 'Taraxacum officinale*')] <- 'Taraxacum officinale'
raw.bloom.ds$plant[which(raw.bloom.ds$plant == 'Thin yellow Brassica ')] <- 'Thin yellow Brassica'
raw.bloom.ds$plant[which(raw.bloom.ds$plant == 'thin yellow Brassica ')] <- 'Thin yellow Brassica'
raw.bloom.ds$plant[which(raw.bloom.ds$plant == 'Viola  adunca')] <- 'Viola adunca'
raw.bloom.ds$plant[which(raw.bloom.ds$plant == 'Viola  praemorsa')] <- 'Viola praemorsa'
raw.bloom.ds <- raw.bloom.ds[which(raw.bloom.ds$plant != 'White aster'), ] 
raw.bloom.ds <- raw.bloom.ds[which(raw.bloom.ds$plant != 'Unknown'), ] 
raw.bloom.ds <- raw.bloom.ds[which(raw.bloom.ds$plant != '"Wallflower"'), ] 

raw.bloom.ds$betterdate <- as.Date(raw.bloom.ds$date, format = "%m/%d/%y")

```

Now you can begin the analysis in the `RMBL-NICHE-Niche-Dyanmics-Analysis.Rmd` using the objects `raw.bloom.ds` and `raw.ds1`

